Tackling irreproducible pathway enrichment analysis

### 2024-03-14 manuscript/manuscript.Rmd
ms workl

---
title: "Tackling irreproducible enrichment analysis"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
bibliography: references.bib
csl: plos-computational-biology.csl
---

Alt: *Tackling irreproducible pathway enrichment analysis*

@@

Irreproducible research is a significant problem that wastes resources and risks
undermining the public perception of science.
Previous work has highlighted that published enrichment analyses frequently suffer from statistical
and reporting flaws.
We sought to determine whether this translates into irreproducibility by examining whether the
findings of 20 open-access articles published in 2019 describing enrichment analysis with the
popular DAVID suite were reproducible.
We find that only four articles exhibited a high degree of concordance, while seven exhibited major
discrepancies.
These inconsistencies can mostly be ascribed to deficiencies in methodological reporting and lack of
preservation of past DAVID versions.
Better reporting and a long-term view of reproducibility are needed.
Scripted workflows and shared code are a potential solution to this problem, although there is a
steep learning curve for novice analysts.
We provide a step-by-step beginner-friendly enrichment analysis protocol
exemplifying current best practices including literate programming, version control, containerisation,
documentation and persistent sharing of data and code.

@@

polygon(c(2,10.5,10.5,2),c(2.5,2.5,1.5,1.5), col=cols2[5], border = NA)
text( 6.3, 2, lab="5. Customize R Markdown: 3-12 hrs",cex=1.5)

@@

>>>>

---
title: "Tackling irreproducible pathway enrichment analysis"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
bibliography: references.bib
csl: plos-computational-biology.csl
---

@@

Irreproducible research is a significant problem that wastes resources and risks
undermining the public perception of science.
Previous work has highlighted that published pathway enrichment analyses frequently suffer from
statistical and reporting flaws.
We sought to determine whether this translates into irreproducibility by examining whether the
findings of 20 open-access articles published in 2019 describing enrichment analysis with the
popular DAVID suite were reproducible.
We find that only four articles exhibited a high degree of concordance, while seven exhibited major
discrepancies.
These inconsistencies can mostly be ascribed to deficiencies in methodological reporting and lack of
preservation of past DAVID versions.
Better reporting and a long-term view of reproducibility are needed.
Scripted workflows and shared code are a potential solution to this problem, although there is a
steep learning curve for novice analysts.
We provide a step-by-step beginner-friendly enrichment analysis protocol
exemplifying current best practices including literate programming, version control, containerisation,
documentation and persistent sharing of data and code.

@@

polygon(c(2,10.5,10.5,2),c(2.5,2.5,1.5,1.5), col=cols2[5], border = NA)
text( 6.3, 2, lab="5. Customise R Markdown: 3-12 hrs",cex=1.5)

@@

Furthermore, our reproducibility assessment is based on "head-counting" pathways that meet an arbitrary significance threshold.
This binary classification of results is less powerful as opposed to comparison of standardised effect sizes.
